### 2018-02-20 test.tex

\documentclass{article}
\usepackage[margin = 0.1cm]{geometry}
\usepackage{tipa}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{longtable}
\begin{document}

\emph{Kikurai:}\\\begin{longtable}{ll}\toprule
Surface form & UR
\\ \midrule
\textipa{aBa\'{a}nto}&\textipa{aBa\'{a}nto}\\
\textipa{aBam\'{u}ra}&\textipa{aBam\'{u}ra}\\
\textipa{amah\'{i}\'{i}ndi}&\textipa{amah\'{i}\'{i}nri}\\
\textipa{amak\'{E}\'{E}ndO}&\textipa{amak\'{E}\'{E}nrO}\\
\textipa{eB\|x{a}}&\textipa{eB\|x{a}}\\
\textipa{eeNgw\'{e}}&\textipa{eeNGw\'{e}}\\
\textipa{eG\|x{a}}&\textipa{eG\|x{a}}\\
\textipa{ekeBw\'{E}}&\textipa{ekeBw\'{E}}\\
\textipa{hoor\'{a}}&\textipa{hoor\'{a}}\\
\textipa{iBiG\'{u}r\'{u}Be}&\textipa{iBiG\'{u}r\'{u}Be}\\
\textipa{iBir\'{u}\'{u}Ng\'{u}uri}&\textipa{iBir\'{u}\'{u}NG\'{u}uri}\\
\textipa{uGus\'{i}ri}&\textipa{uGus\'{i}ri}\\
\textipa{B\'{a}inu}&\textipa{B\'{a}inu}\\
\textipa{Borj\'{o}}&\textipa{Borj\'{o}}\\
\textipa{it\super SiiNg\'{E}na}&\textipa{it\super SiiNG\'{E}na}\\
\textipa{it\super SiiNg\'{u}r\'{u}Be}&\textipa{it\super SiiNG\'{u}r\'{u}Be}\\
\textipa{GaB\|x{a}}&\textipa{GaB\|x{a}}\\
\textipa{it\super SiiNg\'{u}ta}&\textipa{it\super SiiNG\'{u}ta}\\
\textipa{BErEk\'{a}}&\textipa{BErEk\'{a}}\\
\textipa{iGit\'{u}\'{u}mbe}&\textipa{iGit\'{u}\'{u}mBe}\\
\textipa{G\'{u}\'{u}k\'{a}}&\textipa{G\'{u}\'{u}k\'{a}}\\
\textipa{rem\|x{a}}&\textipa{rem\|x{a}}\\
\textipa{rEEnt\'{a}}&\textipa{rEEnt\'{a}}\\
\textipa{oBoG\'{a}\'{a}k\'{a}}&\textipa{oBoG\'{a}\'{a}k\'{a}}\\
\textipa{oBot\'{e}\'{e}nd\'{e}\'{e}ru}&\textipa{oBot\'{e}\'{e}nr\'{e}\'{e}ru}\\
\textipa{okoG\'{e}\'{e}mb\'{a}}&\textipa{okoG\'{e}\'{e}mB\'{a}}\\
\textipa{oko\'{o}mb\'{a}ra}&\textipa{oko\'{o}mB\'{a}ra}\\
\textipa{okoB\'{a}ra}&\textipa{okoB\'{a}ra}\\
\textipa{oko\'{o}nd\'{O}Ga}&\textipa{oko\'{o}nr\'{O}Ga}\\
\textipa{okor\'{O}Ga}&\textipa{okor\'{O}Ga}\\
\textipa{rom\|x{a}}&\textipa{rom\|x{a}}\\
\textipa{teGet\'{a}}&\textipa{teGet\'{a}}\\
\textipa{uku\'{u}mbuurj\'{a}}&\textipa{uku\'{u}mBuurj\'{a}}\\
\textipa{uruG\'{u}ta}&\textipa{uruG\'{u}ta}\\
\bottomrule\end{longtable}

\begin{longtable}{ll}\toprule
\emph{The surface form...}&\emph{Is underlyingly...}
\\ \midrule
\textipa{b}&\textipa{B}\\
\textipa{d}&\textipa{r}\\
\textipa{g}&\textipa{G}\\
\bottomrule\end{longtable}

\begin{tabular}{l}\emph{Rules: }\\
\verb|[+liquid]| $\to$ \textipa{d} / \verb|[-vowel]| \verb|_| \\\verb|[-sonorant]| $\to$ \verb|[-continuant]| / \verb|[-vowel]| \verb|_| 
\end{tabular}

\pagebreak

\emph{Modern Greek:}\\\begin{longtable}{ll}\toprule
Surface form & UR
\\ \midrule
\textipa{kano}&\textipa{kano}\\
\textipa{kori}&\textipa{kori}\\
\textipa{xano}&\textipa{xano}\\
\textipa{xori}&\textipa{xori}\\
\textipa{x\super jino}&\textipa{xino}\\
\textipa{k\super jino}&\textipa{kino}\\
\textipa{krima}&\textipa{krima}\\
\textipa{xrima}&\textipa{xrima}\\
\textipa{xufta}&\textipa{xufta}\\
\textipa{kufeta}&\textipa{kufeta}\\
\textipa{kali}&\textipa{kali}\\
\textipa{xali}&\textipa{xali}\\
\textipa{x\super jeli}&\textipa{xeli}\\
\textipa{k\super jeri}&\textipa{keri}\\
\textipa{x\super jeri}&\textipa{xeri}\\
\textipa{ox\super ji}&\textipa{oxi}\\
\bottomrule\end{longtable}

\begin{longtable}{ll}\toprule
\emph{The surface form...}&\emph{Is underlyingly...}
\\ \midrule
\textipa{x\super j}&\textipa{x}\\
\textipa{k\super j}&\textipa{k}\\
\bottomrule\end{longtable}

\begin{tabular}{l}\emph{Rules: }\\
\verb|[+back -voice]| $\to$ \verb|[+palletized]| /  \verb|_| \verb|[-liquid -back]|
\end{tabular}

\pagebreak

\emph{Farsi:}\\\begin{longtable}{ll}\toprule
Surface form & UR
\\ \midrule
\textipa{\ae \~rteS}&\textipa{\ae \~rteS}\\
\textipa{fa\~rsi}&\textipa{fa\~rsi}\\
\textipa{q\ae d\~ri}&\textipa{q\ae d\~ri}\\
\textipa{\~rah}&\textipa{\~rah}\\
\textipa{\~rast}&\textipa{\~rast}\\
\textipa{\~riS}&\textipa{\~riS}\\
\textipa{aha\r*{\~r}}&\textipa{aha\r*{\~r}}\\
\textipa{ax\ae \r*{\~r}}&\textipa{ax\ae \r*{\~r}}\\
\textipa{h\ae \~rtow\r*{\~r}}&\textipa{h\ae \~rtow\r*{\~r}}\\
\textipa{Si\r*{\~r}}&\textipa{Si\r*{\~r}}\\
\textipa{aha\v{r}i}&\textipa{aha\~ri}\\
\textipa{b\ae \v{r}ad\ae \r*{\~r}}&\textipa{b\ae \~rad\ae \r*{\~r}}\\
\textipa{t\super Se\v{r}a}&\textipa{t\super Se\~ra}\\
\textipa{da\v{r}id}&\textipa{da\~rid}\\
\textipa{bi\v{r}\ae ng}&\textipa{bi\~r\ae ng}\\
\textipa{Si\v{r}ini}&\textipa{Si\~rini}\\
\bottomrule\end{longtable}

\begin{longtable}{ll}\toprule
\emph{The surface form...}&\emph{Is underlyingly...}
\\ \midrule
\textipa{\v{r}}&\textipa{\~r}\\
\bottomrule\end{longtable}

\begin{tabular}{l}\emph{Rules: }\\
\verb|[+trill]| $\to$ \textipa{\v{r}} / \verb|[+vowel]| \verb|_| \verb|[+voice]|
\end{tabular}

\pagebreak

\emph{Osage:}\\\begin{longtable}{ll}\toprule
Surface form & UR
\\ \midrule
\textipa{d\'{a}br\~i}&\textipa{d\'{a}br\~i}\\
\textipa{\'{a}Dik\super h\~aZ\~a}&\textipa{\'{a}dik\super h\~aZ\~a}\\
\textipa{dat\super Sp\'{e}}&\textipa{dat\super Sp\'{e}}\\
\textipa{t\super SP\'{e}De}&\textipa{t\super SP\'{e}de}\\
\textipa{dakP\'{e}}&\textipa{dakP\'{e}}\\
\textipa{D\'{e}ze}&\textipa{d\'{e}ze}\\
\textipa{d\'{a}l\~i}&\textipa{d\'{a}l\~i}\\
\textipa{D\'{i}e}&\textipa{d\'{i}e}\\
\textipa{daSt\'{u}}&\textipa{daSt\'{u}}\\
\textipa{D\'{i}Ski}&\textipa{d\'{i}Ski}\\
\bottomrule\end{longtable}

\begin{longtable}{ll}\toprule
\emph{The surface form...}&\emph{Is underlyingly...}
\\ \midrule
\textipa{D}&\textipa{d}\\
\bottomrule\end{longtable}

\begin{tabular}{l}\emph{Rules: }\\
\textipa{d} $\to$ \textipa{D} /  \verb|_| \verb|[-low]|
\end{tabular}

\pagebreak

\emph{Amharic:}\\\begin{longtable}{ll}\toprule
Surface form & UR
\\ \midrule
\textipa{f@r@s}&\textipa{f@r@s}\\
\textipa{t@n@sa}&\textipa{t@n@sa}\\
\textipa{jEl1Sl1S}&\textipa{j@l1Sl1S}\\
\textipa{majEt}&\textipa{maj@t}\\
\textipa{g@nz@b}&\textipa{g@nz@b}\\
\textipa{SEgna}&\textipa{S@gna}\\
\textipa{n@\~n}&\textipa{n@\~n}\\
\textipa{m@wd@d}&\textipa{m@wd@d}\\
\textipa{m@nn@sat}&\textipa{m@nn@sat}\\
\textipa{m@mk@r}&\textipa{m@mk@r}\\
\textipa{ZEle}&\textipa{Z@le}\\
\textipa{jEll@m}&\textipa{j@ll@m}\\
\textipa{m@t\super S}&\textipa{m@t\super S}\\
\textipa{m@st@t}&\textipa{m@st@t}\\
\textipa{f@ll@g@}&\textipa{f@ll@g@}\\
\textipa{ag@\~n\~nE}&\textipa{ag@\~n\~n@}\\
\textipa{t@m@t\super St\super SE}&\textipa{t@m@t\super St\super S@}\\
\textipa{mokk@r@}&\textipa{mokk@r@}\\
\textipa{kaZZE}&\textipa{kaZZ@}\\
\textipa{ZEmm@r@}&\textipa{Z@mm@r@}\\
\textipa{lat\super St\super SE}&\textipa{lat\super St\super S@}\\
\textipa{aSSE}&\textipa{aSS@}\\
\textipa{b@kk@l@}&\textipa{b@kk@l@}\\
\textipa{SEm@gg@l@}&\textipa{S@m@gg@l@}\\
\bottomrule\end{longtable}

\begin{longtable}{ll}\toprule
\emph{The surface form...}&\emph{Is underlyingly...}
\\ \midrule
\textipa{E}&\textipa{@}\\
\bottomrule\end{longtable}

\begin{tabular}{l}\emph{Rules: }\\
\textipa{@} $\to$ \textipa{E} / \verb|[-anterior -liquid -high]| \verb|_| 
\end{tabular}

\pagebreak

\emph{Gen:}\\\begin{longtable}{ll}\toprule
Surface form & UR
\\ \midrule
\textipa{agble}&\textipa{agble}\\
\textipa{agoNglo}&\textipa{agoNglo}\\
\textipa{aNOli}&\textipa{aNOli}\\
\textipa{akplO}&\textipa{akplO}\\
\textipa{sabulE}&\textipa{sabulE}\\
\textipa{sra}&\textipa{sla}\\
\textipa{alO}&\textipa{alO}\\
\textipa{atitrwE}&\textipa{atitlwE}\\
\textipa{avlO}&\textipa{avlO}\\
\textipa{blafogbe}&\textipa{blafogbe}\\
\textipa{drE}&\textipa{dlE}\\
\textipa{edrO}&\textipa{edlO}\\
\textipa{exlO}&\textipa{exlO}\\
\textipa{exle}&\textipa{exle}\\
\textipa{hlE}&\textipa{hlE}\\
\textipa{NlO}&\textipa{NlO}\\
\textipa{t\super Sr\~O}&\textipa{t\super Sl\~O}\\
\textipa{\~nr\~a}&\textipa{\~nl\~a}\\
\textipa{klO}&\textipa{klO}\\
\textipa{tre}&\textipa{tle}\\
\textipa{vlu}&\textipa{vlu}\\
\textipa{lO}&\textipa{lO}\\
\textipa{mla}&\textipa{mla}\\
\textipa{pleplelu}&\textipa{pleplelu}\\
\textipa{wla}&\textipa{wla}\\
\textipa{zro}&\textipa{zlo}\\
\textipa{esrO}&\textipa{eslO}\\
\textipa{etro}&\textipa{etlo}\\
\textipa{e\~nr\~O}&\textipa{e\~nl\~O}\\
\textipa{d\super Zro}&\textipa{d\super Zlo}\\
\bottomrule\end{longtable}

\begin{longtable}{ll}\toprule
\emph{The surface form...}&\emph{Is underlyingly...}
\\ \midrule
\textipa{r}&\textipa{l}\\
\bottomrule\end{longtable}

\begin{tabular}{l}\emph{Rules: }\\
\verb|[ ]| $\to$ \verb|[-lateral]| / \verb|[+coronal]| \verb|_| 
\end{tabular}

\pagebreak

\emph{Kishambaa:}\\\begin{longtable}{ll}\toprule
Surface form & UR
\\ \midrule
\textipa{tagi}&\textipa{tagi}\\
\textipa{kitabu}&\textipa{kitabu}\\
\textipa{paalika}&\textipa{paalika}\\
\textipa{ni}&\textipa{ni}\\
\textipa{Nombe}&\textipa{Nombe}\\
\textipa{matagi}&\textipa{matagi}\\
\textipa{dodoa}&\textipa{dodoa}\\
\textipa{goSa}&\textipa{goSa}\\
\textipa{babu}&\textipa{babu}\\
\textipa{ndimi}&\textipa{ndimi}\\
\textipa{Ngoto}&\textipa{Ngoto}\\
\textipa{mbeu}&\textipa{mbeu}\\
\textipa{\r*nt\super humbii}&\textipa{nt\super humbii}\\
\textipa{Nok\super huNguni}&\textipa{Nok\super huNguni}\\
\textipa{\r*mp\super heho}&\textipa{mp\super heho}\\
\bottomrule\end{longtable}

\begin{longtable}{ll}\toprule
\emph{The surface form...}&\emph{Is underlyingly...}
\\ \midrule
\textipa{\r*m}&\textipa{m}\\
\textipa{\r*n}&\textipa{n}\\
\bottomrule\end{longtable}

\begin{tabular}{l}\emph{Rules: }\\
\verb|[+nasal]| $\to$ \verb|[-voice]| /  \verb|_| \verb|[-voice]|
\end{tabular}

\pagebreak

\emph{Thai:}\\\begin{longtable}{ll}\toprule
Surface form & UR
\\ \midrule
\textipa{bil}&\textipa{bil}\\
\textipa{m\"u\"u}&\textipa{m\"u\"u}\\
\textipa{rak\textcorner }&\textipa{rak}\\
\textipa{baa}&\textipa{baa}\\
\textipa{loN}&\textipa{loN}\\
\textipa{br\"u\"u}&\textipa{br\"u\"u}\\
\textipa{haa}&\textipa{haa}\\
\textipa{plaa}&\textipa{plaa}\\
\textipa{dii}&\textipa{dii}\\
\textipa{t\super Saan}&\textipa{t\super Saan}\\
\textipa{t\super hee}&\textipa{t\super hee}\\
\textipa{t\super hruumEEn}&\textipa{t\super hruumEEn}\\
\textipa{k\super hEN}&\textipa{k\super hEN}\\
\textipa{panjaa}&\textipa{panjaa}\\
\textipa{l@@j}&\textipa{l@@j}\\
\textipa{p\super hjaa}&\textipa{p\super hjaa}\\
\textipa{l\"uak\textcorner }&\textipa{l\"uak}\\
\textipa{klaaN}&\textipa{klaaN}\\
\textipa{t\super S\super hat\textcorner }&\textipa{t\super S\super hat}\\
\textipa{traa}&\textipa{traa}\\
\textipa{riip\textcorner }&\textipa{riip}\\
\textipa{OOk\textcorner }&\textipa{OOk}\\
\textipa{p\super hrEE}&\textipa{p\super hrEE}\\
\textipa{ki@}&\textipa{ki@}\\
\textipa{k\super hwaa}&\textipa{k\super hwaa}\\
\textipa{kEE}&\textipa{kEE}\\
\textipa{draj}&\textipa{draj}\\
\textipa{d\"uN}&\textipa{d\"uN}\\
\textipa{kan}&\textipa{kan}\\
\textipa{t\super Su@k\textcorner }&\textipa{t\super Su@k}\\
\textipa{p\super hleeN}&\textipa{p\super hleeN}\\
\textipa{t\super S\super han}&\textipa{t\super S\super han}\\
\textipa{staaN}&\textipa{staaN}\\
\textipa{rap\textcorner }&\textipa{rap}\\
\textipa{jiisip\textcorner }&\textipa{jiisip}\\
\textipa{p\super haa}&\textipa{p\super haa}\\
\textipa{k\super haa}&\textipa{k\super haa}\\
\textipa{dam}&\textipa{dam}\\
\textipa{raaj}&\textipa{raaj}\\
\textipa{tit\textcorner }&\textipa{tit}\\
\textipa{sip\textcorner }&\textipa{sip}\\
\textipa{pen}&\textipa{pen}\\
\bottomrule\end{longtable}

\begin{longtable}{ll}\toprule
\emph{The surface form...}&\emph{Is underlyingly...}
\\ \midrule
\textipa{t\textcorner }&\textipa{t}\\
\textipa{p\textcorner }&\textipa{p}\\
\textipa{k\textcorner }&\textipa{k}\\
\bottomrule\end{longtable}

\begin{tabular}{l}\emph{Rules: }\\
\verb|[-voice]| $\to$ \verb|[+unreleased]| /  \verb|_| \#
\end{tabular}

\pagebreak

\emph{Palauan:}\\\begin{longtable}{ll}\toprule
Surface form & UR
\\ \midrule
\textipa{k@D@}&\textipa{k@D@}\\
\textipa{b@Duk}&\textipa{b@Duk}\\
\textipa{Diak}&\textipa{Diak}\\
\textipa{diak}&\textipa{diak}\\
\textipa{maTtNoT}&\textipa{madtNod}\\
\textipa{De:l}&\textipa{De:l}\\
\textipa{de:l}&\textipa{de:l}\\
\textipa{Dios@P}&\textipa{Dios@P}\\
\textipa{dios@P}&\textipa{dios@P}\\
\textipa{Dik}&\textipa{Dik}\\
\textipa{dik}&\textipa{dik}\\
\textipa{kuT}&\textipa{kud}\\
\textipa{PoDiN@l}&\textipa{PoDiN@l}\\
\textipa{koaT}&\textipa{koad}\\
\textipa{eaN@T}&\textipa{eaN@d}\\
\textipa{N@rar@D@}&\textipa{N@rar@D@}\\
\textipa{baT}&\textipa{bad}\\
\textipa{ieDlP@Dip}&\textipa{ieDlP@Dip}\\
\textipa{k@Deb}&\textipa{k@Deb}\\
\textipa{m@D@Nei}&\textipa{m@D@Nei}\\
\textipa{uDouT}&\textipa{uDoud}\\
\textipa{olDak}&\textipa{olDak}\\
\bottomrule\end{longtable}

\begin{longtable}{ll}\toprule
\emph{The surface form...}&\emph{Is underlyingly...}
\\ \midrule
\textipa{T}&\textipa{d}\\
\bottomrule\end{longtable}

\begin{tabular}{l}\emph{Rules: }\\
\textipa{d} $\to$ \textipa{T} / \verb|[ ]| \verb|_| 
\end{tabular}

\pagebreak

\emph{Quechua (Cuzco dialect):}\\\begin{longtable}{ll}\toprule
Surface form & UR
\\ \midrule
\textipa{qori}&\textipa{quri}\\
\textipa{t\super SoXlu}&\textipa{t\super SuXlu}\\
\textipa{qomir}&\textipa{qumir}\\
\textipa{niNri}&\textipa{niNri}\\
\textipa{moqo}&\textipa{muqu}\\
\textipa{hoqara}&\textipa{huqara}\\
\textipa{p\super hulju}&\textipa{p\super hulju}\\
\textipa{jujaN}&\textipa{jujaN}\\
\textipa{tulju}&\textipa{tulju}\\
\textipa{api}&\textipa{api}\\
\textipa{suti}&\textipa{suti}\\
\textipa{o\;Nqoj}&\textipa{uNquj}\\
\textipa{t\super Silwi}&\textipa{t\super Silwi}\\
\textipa{t\super S\super hit\super SiN}&\textipa{t\super S\super hit\super SiN}\\
\textipa{t\super S\super ha\;Nqaj}&\textipa{t\super S\super haNqaj}\\
\textipa{a\;Nqosaj}&\textipa{aNqusaj}\\
\textipa{qet\super SuN}&\textipa{qit\super SuN}\\
\textipa{pisqo}&\textipa{pisqu}\\
\textipa{musoX}&\textipa{musuX}\\
\textipa{t\super SuNka}&\textipa{t\super SuNka}\\
\textipa{ja\;NqaN}&\textipa{jaNqaN}\\
\textipa{t\super Sulju}&\textipa{t\super Sulju}\\
\textipa{qhelja}&\textipa{qhilja}\\
\textipa{qe\;Nqo}&\textipa{qiNqu}\\
\textipa{t\super SeqaN}&\textipa{t\super SiqaN}\\
\textipa{qaN}&\textipa{qaN}\\
\textipa{noqa}&\textipa{nuqa}\\
\textipa{t\super Saxra}&\textipa{t\super Saxra}\\
\textipa{t\super SeXniN}&\textipa{t\super SiXniN}\\
\textipa{soXta}&\textipa{suXta}\\
\textipa{aXna}&\textipa{aXna}\\
\textipa{ljixlja}&\textipa{ljixlja}\\
\textipa{qosa}&\textipa{qusa}\\
\textipa{qara}&\textipa{qara}\\
\textipa{alqo}&\textipa{alqu}\\
\textipa{se\;Nqa}&\textipa{siNqa}\\
\textipa{karu}&\textipa{karu}\\
\textipa{atoX}&\textipa{atuX}\\
\textipa{qaNkuna}&\textipa{qaNkuna}\\
\textipa{pusaX}&\textipa{pusaX}\\
\textipa{teXwaj}&\textipa{tiXwaj}\\
\textipa{t\super Saki}&\textipa{t\super Saki}\\
\textipa{wateX}&\textipa{watiX}\\
\textipa{aNka}&\textipa{aNka}\\
\textipa{waXtaj}&\textipa{waXtaj}\\
\textipa{haku}&\textipa{haku}\\
\textipa{waqaj}&\textipa{waqaj}\\
\textipa{kaNka}&\textipa{kaNka}\\
\textipa{waxt\super Sa}&\textipa{waxt\super Sa}\\
\textipa{waleX}&\textipa{waliX}\\
\textipa{t\super hakaj}&\textipa{t\super hakaj}\\
\textipa{reXsisqa}&\textipa{riXsisqa}\\
\bottomrule\end{longtable}

\begin{longtable}{ll}\toprule
\emph{The surface form...}&\emph{Is underlyingly...}
\\ \midrule
\textipa{e}&\textipa{i}\\
\textipa{o}&\textipa{u}\\
\textipa{\;N}&\textipa{N}\\
\bottomrule\end{longtable}

\begin{tabular}{l}\emph{Rules: }\\
\verb|[+high]| $\to$ \textipa{\;N} /  \verb|_| \textipa{q}\\\verb|[ ]| $\to$ \verb|[-high]| / \textipa{q} \verb|_| \\\textipa{u} $\to$ \textipa{o} /  \verb|_| \verb|[+back -high]|\\\verb|[+sonorant]| $\to$ \verb|[-high]| / \verb|[-palletized]| \verb|_| \verb|[-anterior]|* \verb|[-nasal +back]|
\end{tabular}

\pagebreak

\emph{Lhasa Tibetan:}\\\begin{longtable}{ll}\toprule
Surface form & UR
\\ \midrule
\textipa{aNgu}&\textipa{aNku}\\
\textipa{aN\.*t\~a\~a}&\textipa{aN\.*t\~a\~a}\\
\textipa{aNba}&\textipa{aNpa}\\
\textipa{apsoo}&\textipa{apsoo}\\
\textipa{amt\super SOO}&\textipa{amt\super SOO}\\
\textipa{tuk\.*t\"u\"u}&\textipa{tuk\.*t\"u\"u}\\
\textipa{amto}&\textipa{amto}\\
\textipa{1Gu}&\textipa{1Gu}\\
\textipa{1mt\super Si}&\textipa{1mt\super Si}\\
\textipa{u\.*t1}&\textipa{u\.*t1}\\
\textipa{uB\textiota \textiota }&\textipa{uB\textiota \textiota }\\
\textipa{ea}&\textipa{ea}\\
\textipa{embo}&\textipa{empo}\\
\textipa{UUtsi}&\textipa{UUtsi}\\
\textipa{qa}&\textipa{ka}\\
\textipa{qaa}&\textipa{kaa}\\
\textipa{qaNba}&\textipa{kaNpa}\\
\textipa{qamba}&\textipa{kampa}\\
\textipa{qam}&\textipa{kam}\\
\textipa{qamtoo}&\textipa{kamtoo}\\
\textipa{qaaBo}&\textipa{kaaBo}\\
\textipa{k1k\.*ti}&\textipa{k1k\.*ti}\\
\textipa{k1Bu}&\textipa{k1Bu}\\
\textipa{k11Nguu}&\textipa{k11Nkuu}\\
\textipa{kik}&\textipa{kik}\\
\textipa{ki\.*tuu}&\textipa{ki\.*tuu}\\
\textipa{k\textiota \textiota cuu}&\textipa{k\textiota \textiota cuu}\\
\textipa{k\textiota \textiota rii}&\textipa{k\textiota \textiota rii}\\
\textipa{kiiGuu}&\textipa{kiiGuu}\\
\textipa{ku}&\textipa{ku}\\
\textipa{kupc1}&\textipa{kupc1}\\
\textipa{kupcaa}&\textipa{kupcaa}\\
\textipa{kEnca}&\textipa{kEnca}\\
\textipa{kEmbo}&\textipa{kEmpo}\\
\textipa{keG\"o\"o}&\textipa{keG\"o\"o}\\
\textipa{kerBa}&\textipa{kerBa}\\
\textipa{qo}&\textipa{ko}\\
\textipa{qomba}&\textipa{kompa}\\
\textipa{qOr}&\textipa{kOr}\\
\textipa{qOOOO}&\textipa{kOOOO}\\
\textipa{t\super S\super hea}&\textipa{t\super S\super hea}\\
\textipa{t\super S\super huGum}&\textipa{t\super S\super huGum}\\
\textipa{topcaa}&\textipa{topcaa}\\
\textipa{t\super ho\~o\~o}&\textipa{t\super ho\~o\~o}\\
\textipa{\.*taa\~a\~a}&\textipa{\.*taa\~a\~a}\\
\textipa{\.*tuG1}&\textipa{\.*tuG1}\\
\textipa{\.*tu\;N\;Go}&\textipa{\.*tuNgo}\\
\textipa{nE\;N\;Gaa}&\textipa{nENgaa}\\
\textipa{pa\;N\;GOO}&\textipa{paNgOO}\\
\textipa{pEEB\~a\~a}&\textipa{pEEB\~a\~a}\\
\textipa{sim\;G\~a\~a}&\textipa{simg\~a\~a}\\
\bottomrule\end{longtable}

\begin{longtable}{ll}\toprule
\emph{The surface form...}&\emph{Is underlyingly...}
\\ \midrule
\textipa{b}&\textipa{p}\\
\textipa{d}&\textipa{t}\\
\textipa{\;G}&\textipa{g}\\
\textipa{\.*d}&\textipa{\.*t}\\
\textipa{\;N}&\textipa{N}\\
\textipa{q}&\textipa{k}\\
\textipa{g}&\textipa{k}\\
\bottomrule\end{longtable}

\begin{tabular}{l}\emph{Rules: }\\
\verb|[-anterior -vowel]| $\to$ \verb|[-sonorant -high]| /  \verb|_| \verb|[+high]|* \verb|[+back -high]|\\\verb|[-coronal]| $\to$ \verb|[+voice]| / \verb|[-vowel]| \verb|_| 
\end{tabular}

\pagebreak

\emph{Axininca Campa:}\\\begin{longtable}{ll|l}\toprule
$\varnothing$ $+$stem$+$ $\varnothing$ & \textipa{no} $+$stem$+$ \textipa{ti} & UR
\\ \midrule
\textipa{toniro} & \textipa{notoniroti} & \textipa{toniro}\\
\textipa{jaarato} & \textipa{nojaaratoti} & \textipa{jaarato}\\
\textipa{kanari} & \textipa{nojanariti} & \textipa{kanari}\\
\textipa{kosiri} & \textipa{nojosiriti} & \textipa{kosiri}\\
\textipa{pisiro} & \textipa{nowisiroti} & \textipa{wisiro}\\
\textipa{porita} & \textipa{noworitati} & \textipa{worita}\\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\textipa{w} $\to$ \textipa{p} / \# \verb|_| \\\textipa{k} $\to$ \textipa{j} / \verb|[ ]| \verb|_| 
\end{tabular}

\pagebreak

\emph{Kikuyu:}\\\begin{longtable}{l|l}\toprule
\textipa{ko} $+$stem$+$ \textipa{a} & UR
\\ \midrule
\textipa{GotENEra} & \textipa{tENEr}\\
\textipa{Gokuua} & \textipa{kuu}\\
\textipa{Gokoora} & \textipa{koor}\\
\textipa{koruGa} & \textipa{ruG}\\
\textipa{kooria} & \textipa{ori}\\
\textipa{komE\~na} & \textipa{mE\~n}\\
\textipa{kohOta} & \textipa{hOt}\\
\textipa{Got\super Sina} & \textipa{t\super Sin}\\
\textipa{koGeera} & \textipa{Geer}\\
\textipa{koina} & \textipa{in}\\
\textipa{Got\super Suuka} & \textipa{t\super Suuk}\\
\textipa{Gokaja} & \textipa{kaj}\\
\textipa{koGaja} & \textipa{Gaj}\\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[ ]| $\to$ \textipa{G} / \# \verb|_| \verb|[ ]| \verb|[-low -voice]|
\end{tabular}

\pagebreak

\emph{Korean:}\\\begin{longtable}{ll|l}\toprule
$\varnothing$ $+$stem$+$ \textipa{a} & $\varnothing$ $+$stem$+$ \textipa{ko} & UR
\\ \midrule
\textipa{ip@} & \textipa{ipko} & \textipa{ip}\\
\textipa{kup@} & \textipa{kupko} & \textipa{kup}\\
\textipa{kap\super ha} & \textipa{kapko} & \textipa{kap\super h}\\
\textipa{cip\super h@} & \textipa{cipko} & \textipa{cip\super h}\\
\textipa{tata} & \textipa{tatko} & \textipa{tat}\\
\textipa{put\super h@} & \textipa{putko} & \textipa{put\super h}\\
\textipa{m@k@} & \textipa{m@kko} & \textipa{m@k}\\
\textipa{t\super Suk@} & \textipa{t\super Sukko} & \textipa{t\super Suk}\\
\textipa{ik@} & \textipa{ikko} & \textipa{ik}\\
\textipa{taka} & \textipa{takko} &  -- \\
\textipa{kaka} & \textipa{kakko} &  -- \\
\textipa{s@k@} & \textipa{s@kko} &  -- \\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[ ]| $\to$ \verb|[-aspirated]| /  \verb|_| \verb|[-voice]|\\\textipa{a} $\to$ \textipa{@} / \verb|[-low]| \verb|[ ]| \verb|_| 
\end{tabular}

\pagebreak

\emph{Hungarian:}\\\begin{longtable}{llll|l}\toprule
$\varnothing$ $+$stem$+$ $\varnothing$ & $\varnothing$ $+$stem$+$ \textipa{ban} & $\varnothing$ $+$stem$+$ \textipa{to:l} & $\varnothing$ $+$stem$+$ \textipa{nak} & UR
\\ \midrule
\textipa{kalap} & \textipa{kalabban} & \textipa{kalapto:l} & \textipa{kalapnak} & \textipa{kalap}\\
\textipa{ku:t} & \textipa{ku:dban} & \textipa{ku:tto:l} & \textipa{ku:tnak} & \textipa{ku:t}\\
\textipa{Za:k} & \textipa{Za:gban} & \textipa{Za:kto:l} & \textipa{Za:knak} & \textipa{Za:k}\\
\textipa{re:s} & \textipa{re:zben} & \textipa{re:st\"o:l} & \textipa{re:snek} & \textipa{re:s}\\
\textipa{Sro:f} & \textipa{Sro:vban} & \textipa{Sro:fto:l} & \textipa{Sro:fnak} & \textipa{Sro:f}\\
\textipa{laka:S} & \textipa{laka:Zban} & \textipa{laka:Sto:l} & \textipa{laka:Snak} & \textipa{laka:S}\\
\textipa{ketret\super s} & \textipa{ketred\super zben} & \textipa{ketret\super st\"o:l} & \textipa{ketret\super snek} & \textipa{ketret\super s}\\
\textipa{test} & \textipa{tezdben} & \textipa{testt\"o:l} & \textipa{testnek} & \textipa{tezt}\\
\textipa{rab} & \textipa{rabban} & \textipa{rapto:l} & \textipa{rabnak} & \textipa{rab}\\
\textipa{ka:d} & \textipa{ka:dban} & \textipa{ka:tto:l} & \textipa{ka:dnak} & \textipa{ka:d}\\
\textipa{meleg} & \textipa{melegben} & \textipa{melekt\"o:l} & \textipa{melegnek} & \textipa{meleg}\\
\textipa{vi:z} & \textipa{vi:zben} & \textipa{vi:st\"o:l} & \textipa{vi:znek} & \textipa{vi:z}\\
\textipa{vara:Z} & \textipa{vara:Zban} & \textipa{vara:Sto:l} & \textipa{vara:Znak} &  -- \\
\textipa{a:g\super j} & \textipa{a:g\super jban} & \textipa{a:k\super jto:l} & \textipa{a:g\super jnak} &  -- \\
\textipa{sem} & \textipa{semben} & \textipa{semt\"o:l} & \textipa{semnek} &  -- \\
\textipa{b\"un} & \textipa{b\"unben} & \textipa{b\"unt\"o:l} & \textipa{b\"unnek} &  -- \\
\textipa{toro\~n} & \textipa{toro\~nban} & \textipa{toro\~nto:l} & \textipa{toro\~nnak} &  -- \\
\textipa{fal} & \textipa{falban} & \textipa{falto:l} & \textipa{falnak} &  -- \\
\textipa{\"o:r} & \textipa{\"o:rben} & \textipa{\"o:rt\"o:l} & \textipa{\"o:rnek} &  -- \\
\textipa{sa:j} & \textipa{sa:jban} & \textipa{sa:jto:l} & \textipa{sa:jnak} &  -- \\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[-sonorant]| $\to$ \verb|[+voice]| /  \verb|_| \verb|[-sonorant +voice]|\\\verb|[-sonorant]| $\to$ \verb|[-voice]| /  \verb|_| \verb|[-voice]|\\\verb|[+vowel]| $\to$ \verb|[-back -low]| / \verb|[-back +vowel]| \verb|[ ]|* \verb|_| 
\end{tabular}

\pagebreak

\emph{Kikuria:}\\\begin{longtable}{ll|l}\toprule
$\varnothing$ $+$stem$+$ \textipa{a} & $\varnothing$ $+$stem$+$ \textipa{era} & UR
\\ \midrule
\textipa{suraaNga} & \textipa{suraaNgera} & \textipa{suraaNg}\\
\textipa{taaNgata} & \textipa{taaNgatera} & \textipa{taaNgat}\\
\textipa{baamba} & \textipa{baambera} & \textipa{baamb}\\
\textipa{reenda} & \textipa{reendera} & \textipa{riend}\\
\textipa{rema} & \textipa{remera} & \textipa{rem}\\
\textipa{hoora} & \textipa{hoorera} & \textipa{hoor}\\
\textipa{roma} & \textipa{romera} & \textipa{rom}\\
\textipa{sooka} & \textipa{sookera} & \textipa{sook}\\
\textipa{tat\super Sora} & \textipa{tat\super Sorera} & \textipa{tat\super Sor}\\
\textipa{siika} & \textipa{seekera} & \textipa{siik}\\
\textipa{tiga} & \textipa{tegera} & \textipa{tig}\\
\textipa{ruga} & \textipa{rogera} & \textipa{rug}\\
\textipa{suka} & \textipa{sokera} &  -- \\
\textipa{huuta} & \textipa{hootera} &  -- \\
\textipa{riiNga} & \textipa{reeNgera} &  -- \\
\textipa{siinda} & \textipa{seendera} &  -- \\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[+tense]| $\to$ \verb|[-high]| /  \verb|_| \verb|[-liquid]|* \textipa{e}
\end{tabular}

\pagebreak

\emph{Farsi:}\\\begin{longtable}{ll|l}\toprule
$\varnothing$ $+$stem$+$ $\varnothing$ & $\varnothing$ $+$stem$+$ \textipa{an} & UR
\\ \midrule
\textipa{z\ae n} & \textipa{z\ae nan} & \textipa{z\ae n}\\
\textipa{l\ae b} & \textipa{l\ae ban} & \textipa{l\ae b}\\
\textipa{h\ae sud} & \textipa{h\ae sudan} & \textipa{h\ae sud}\\
\textipa{b\ae rad\ae r} & \textipa{b\ae rad\ae ran} & \textipa{b\ae rad\ae r}\\
\textipa{bozorg} & \textipa{bozorgan} & \textipa{bozorg}\\
\textipa{m\ae leke} & \textipa{m\ae lekean} & \textipa{m\ae leke}\\
\textipa{valede} & \textipa{valedean} & \textipa{valede}\\
\textipa{k\ae bire} & \textipa{k\ae birean} & \textipa{k\ae bire}\\
\textipa{ahu} & \textipa{ahuan} & \textipa{ahu}\\
\textipa{hamele} & \textipa{hamelean} & \textipa{hamele}\\
\textipa{bat\super St\super Se} & \textipa{bat\super St\super Segan} & \textipa{bat\super St\super Seg}\\
\textipa{setare} & \textipa{setaregan} & \textipa{setareg}\\
\textipa{b\ae nde} & \textipa{b\ae ndegan} & \textipa{b\ae ndeg}\\
\textipa{azade} & \textipa{azadegan} &  -- \\
\textipa{divane} & \textipa{divanegan} &  -- \\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[ ]| $\to$ $\varnothing$ / \textipa{e} \verb|_| \#
\end{tabular}

\pagebreak

\emph{Tibetan:}\\\begin{longtable}{ll}\toprule
Number & Surface form
\\ \midrule
10 & \textipa{d\super Zu}\\
1 & \textipa{d\super Zig}\\
11 & \textipa{d\super Zugd\super Zig}\\
4 & \textipa{Si}\\
14 & \textipa{d\super ZubSi}\\
40 & \textipa{Sibd\super Zu}\\
9 & \textipa{gu}\\
19 & \textipa{d\super Zurgu}\\
90 & \textipa{gubd\super Zu}\\
5 & \textipa{Na}\\
15 & \textipa{d\super ZuNa}\\
50 & \textipa{Nabd\super Zu}\\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[-nasal]| $\to$ $\varnothing$ / \# \verb|_| 
\end{tabular}

\pagebreak

\emph{Makonde:}\\\begin{longtable}{lll|l}\toprule
$\varnothing$ $+$stem$+$ \textipa{\'{a}Nga} & $\varnothing$ $+$stem$+$ \textipa{\'{i}le} & $\varnothing$ $+$stem$+$ \textipa{a} & UR
\\ \midrule
\textipa{am\'{a}Nga} & \textipa{am\'{i}le} & \textipa{\'{a}ma} & \textipa{\'{a}m}\\
\textipa{tav\'{a}Nga} & \textipa{tav\'{i}le} & \textipa{t\'{a}va} & \textipa{tav}\\
\textipa{ak\'{a}Nga} & \textipa{ak\'{i}le} & \textipa{\'{a}ka} & \textipa{\'{a}k}\\
\textipa{pat\'{a}Nga} & \textipa{pat\'{i}le} & \textipa{p\'{o}ta} & \textipa{p\'{o}t}\\
\textipa{tat\'{a}Nga} & \textipa{tat\'{i}le} & \textipa{t\'{o}ta} & \textipa{t\'{o}t}\\
\textipa{dab\'{a}Nga} & \textipa{dab\'{i}le} & \textipa{d\'{o}ba} & \textipa{d\'{o}b}\\
\textipa{av\'{a}Nga} & \textipa{av\'{i}le} & \textipa{\'{o}va} & \textipa{\'{o}v}\\
\textipa{am\'{a}Nga} & \textipa{am\'{i}le} & \textipa{\'{o}ma} & \textipa{\'{o}m}\\
\textipa{tap\'{a}Nga} & \textipa{tap\'{i}le} & \textipa{t\'{e}pa} & \textipa{t\'{e}p}\\
\textipa{pat\'{a}Nga} & \textipa{pat\'{i}le} & \textipa{p\'{e}ta} & \textipa{p\'{e}t}\\
\textipa{av\'{a}Nga} & \textipa{av\'{i}le} & \textipa{\'{e}va} & \textipa{\'{e}v}\\
\textipa{bab\'{a}Nga} & \textipa{bab\'{i}le} & \textipa{b\'{e}ba} & \textipa{b\'{e}b}\\
\textipa{ut\'{a}Nga} & \textipa{ut\'{i}le} & \textipa{\'{u}ta} & \textipa{ut}\\
\textipa{luk\'{a}Nga} & \textipa{luk\'{i}le} & \textipa{l\'{u}ka} & \textipa{luk}\\
\textipa{lum\'{a}Nga} & \textipa{lum\'{i}le} & \textipa{l\'{u}ma} & \textipa{lum}\\
\textipa{uNg\'{a}Nga} & \textipa{uNg\'{i}le} & \textipa{\'{u}Nga} & \textipa{uNg}\\
\textipa{iv\'{a}Nga} & \textipa{iv\'{i}le} & \textipa{\'{i}va} & \textipa{iv}\\
\textipa{pit\'{a}Nga} & \textipa{pit\'{i}le} & \textipa{p\'{i}ta} & \textipa{pit}\\
\textipa{imb\'{a}Nga} & \textipa{imb\'{i}le} & \textipa{\'{i}mba} &  -- \\
\textipa{lim\'{a}Nga} & \textipa{lim\'{i}le} & \textipa{l\'{i}ma} &  -- \\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[+vowel]| $\to$ \verb|[+highTone]| /  \verb|_| \verb|[-highTone]|* \verb|[+vowel]| \#\\\verb|[+highTone]| $\to$ \textipa{a} /  \verb|_| \verb|[ ]| \verb|[+highTone]|
\end{tabular}

\pagebreak

\emph{North Saami:}\\\begin{longtable}{ll|l}\toprule
$\varnothing$ $+$stem$+$ $\varnothing$ & $\varnothing$ $+$stem$+$ \textipa{in} & UR
\\ \midrule
\textipa{varit} & \textipa{varihin} & \textipa{varih}\\
\textipa{oahpis} & \textipa{oahpisin} & \textipa{oahpis}\\
\textipa{t\super SoarvuS} & \textipa{t\super SoarvuSin} & \textipa{t\super SoarvuS}\\
\textipa{lottaaS} & \textipa{lottaad\super Zin} & \textipa{lottaad\super Z}\\
\textipa{t\super Suoivvat} & \textipa{t\super Suoivvagin} & \textipa{t\super Suoivvag}\\
\textipa{ahhkut} & \textipa{ahhkubin} & \textipa{ahhkub}\\
\textipa{suohkat} & \textipa{suohkaDin} & \textipa{suohkaD}\\
\textipa{heed\super ZoS} & \textipa{heed\super Zod\super Zin} & \textipa{heed\super Zod\super Z}\\
\textipa{aad\super Zd\super Zut} & \textipa{aad\super Zd\super Zubin} & \textipa{aad\super Zd\super Zub}\\
\textipa{bissobeahtset} & \textipa{bissobeahtsehin} & \textipa{bissobeahtseh}\\
\textipa{t\super Seahtsit} & \textipa{t\super Seahtsibin} & \textipa{t\super Seahtsib}\\
\textipa{jaaPmin} & \textipa{jaaPmimin} & \textipa{jaaPmim}\\
\textipa{t\super Suoivat} & \textipa{t\super Suoivagin} & \textipa{t\super Suoivag}\\
\textipa{laageS} & \textipa{laaged\super Zin} & \textipa{laaged\super Z}\\
\textipa{gahpir} & \textipa{gahpirin} & \textipa{gahpir}\\
\textipa{gaauhtsis} & \textipa{gaauhtsisin} & \textipa{gaauhtsis}\\
\textipa{aaslat} & \textipa{aaslagin} & \textipa{aaslag}\\
\textipa{baDooSgaattset} & \textipa{baDooSgaattsebin} &  -- \\
\textipa{ahhkit} & \textipa{ahhkiDin} &  -- \\
\textipa{bahaanaalat} & \textipa{bahaanaalagin} &  -- \\
\textipa{beStor} & \textipa{beStorin} &  -- \\
\textipa{heevemeahhtun} & \textipa{heevemeahhtunin} &  -- \\
\textipa{beed\super Zot} & \textipa{beed\super Zohin} &  -- \\
\textipa{bissomeahtun} & \textipa{bissomeahtumin} &  -- \\
\textipa{laDas} & \textipa{laDasin} &  -- \\
\textipa{heaijusmielat} & \textipa{heaijusmielagin} &  -- \\
\textipa{heaNkkan} & \textipa{heaNkkanin} &  -- \\
\textipa{jaman} & \textipa{jamanin} &  -- \\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[-liquid -nasal -sibilant]| $\to$ \textipa{t} /  \verb|_| \#\\\verb|[ ]| $\to$ \verb|[+coronal]| /  \verb|_| \#\\\verb|[-sonorant +voice]| $\to$ \textipa{S} /  \verb|_| \#
\end{tabular}

\pagebreak

\emph{Samoan:}\\\begin{longtable}{ll|l}\toprule
$\varnothing$ $+$stem$+$ $\varnothing$ & $\varnothing$ $+$stem$+$ \textipa{ia} & UR
\\ \midrule
\textipa{olo} & \textipa{oloia} & \textipa{olo}\\
\textipa{lafo} & \textipa{lafoia} & \textipa{lafo}\\
\textipa{usu} & \textipa{usuia} & \textipa{usu}\\
\textipa{tau} & \textipa{tauia} & \textipa{tau}\\
\textipa{taui} & \textipa{tauia} & \textipa{taui}\\
\textipa{lele} & \textipa{lelea} & \textipa{lele}\\
\textipa{tafe} & \textipa{tafea} & \textipa{tafe}\\
\textipa{tau} & \textipa{taulia} & \textipa{taul}\\
\textipa{oso} & \textipa{osofia} & \textipa{osof}\\
\textipa{valu} & \textipa{valusia} & \textipa{valus}\\
\textipa{u:} & \textipa{u:tia} &  -- \\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[-tense]| $\to$ $\varnothing$ /  \verb|_| \#\\\verb|[+vowel]| $\to$ $\varnothing$ / \verb|[-back +tense]| \verb|_| \verb|[+vowel]|
\end{tabular}

\pagebreak

\emph{Russian:}\\\begin{longtable}{ll|l}\toprule
$\varnothing$ $+$stem$+$ $\varnothing$ & $\varnothing$ $+$stem$+$ \textipa{a} & UR
\\ \midrule
\textipa{vagon} & \textipa{vagona} & \textipa{vagon}\\
\textipa{glas} & \textipa{glaza} & \textipa{glaz}\\
\textipa{golos} & \textipa{golosa} & \textipa{golos}\\
\textipa{ras} & \textipa{raza} & \textipa{raz}\\
\textipa{les} & \textipa{lesa} & \textipa{les}\\
\textipa{porok} & \textipa{poroga} & \textipa{porog}\\
\textipa{vrak} & \textipa{vraga} & \textipa{vrag}\\
\textipa{urok} & \textipa{uroka} & \textipa{urok}\\
\textipa{tvet} & \textipa{tveta} &  -- \\
\textipa{prut} & \textipa{pruda} &  -- \\
\textipa{soldat} & \textipa{soldata} &  -- \\
\textipa{zavot} & \textipa{zavoda} &  -- \\
\textipa{xlep} & \textipa{xleba} &  -- \\
\textipa{grip} & \textipa{griba} &  -- \\
\textipa{trup} & \textipa{trupa} &  -- \\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[-sonorant]| $\to$ \verb|[-voice]| /  \verb|_| \#
\end{tabular}

\pagebreak

\emph{English:}\\\begin{longtable}{lll|l}\toprule
$\varnothing$ $+$stem$+$ $\varnothing$ & $\varnothing$ $+$stem$+$ \textipa{d} & $\varnothing$ $+$stem$+$ \textipa{@z} & UR
\\ \midrule
\textipa{ro} & \textipa{rod} & \textipa{roz} & \textipa{ro}\\
\textipa{gr\ae b} & \textipa{gr\ae bd} & \textipa{gr\ae bz} & \textipa{gr\ae b}\\
\textipa{sim} & \textipa{simd} & \textipa{simz} & \textipa{sim}\\
\textipa{liv} & \textipa{livd} & \textipa{livz} & \textipa{liv}\\
\textipa{h@g} & \textipa{h@gd} & \textipa{h@gz} & \textipa{h@g}\\
\textipa{lUk} & \textipa{lUkt} & \textipa{lUks} & \textipa{lUk}\\
\textipa{\ae sk} & \textipa{\ae skt} & \textipa{\ae sks} & \textipa{\ae sk}\\
\textipa{wErk} & \textipa{wErkt} & \textipa{wErks} & \textipa{wErk}\\
\textipa{k\textiota s} & \textipa{k\textiota st} & \textipa{k\textiota s@z} & \textipa{k\textiota s}\\
\textipa{f\textiota S} & \textipa{f\textiota St} & \textipa{f\textiota S@z} & \textipa{f\textiota S}\\
\textipa{kw\textiota z} & \textipa{kw\textiota zd} & \textipa{kw\textiota z@z} & \textipa{kw\textiota z}\\
\textipa{b2z} & \textipa{b2zd} & \textipa{b2z@z} & \textipa{b2z}\\
\textipa{wet} & \textipa{wet@d} & \textipa{wets} & \textipa{wet}\\
\textipa{wed} & \textipa{wed@d} & \textipa{wedz} & \textipa{wed}\\
\textipa{l\textiota ft} & \textipa{l\textiota ft@d} & \textipa{l\textiota fts} & \textipa{l\textiota ft}\\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[+vowel]| $\to$ $\varnothing$ / \verb|[+continuant]| \verb|[-sibilant]| \verb|_| \\$\varnothing$ $\to$ \textipa{@} / \verb|[-continuant +coronal]| \verb|_| \verb|[-continuant]|\\\verb|[-sonorant]| $\to$ \verb|[-voice]| / \verb|[-voice]| \verb|_| 
\end{tabular}

\pagebreak

\emph{Finnish:}\\\begin{longtable}{ll|l}\toprule
$\varnothing$ $+$stem$+$ $\varnothing$ & $\varnothing$ $+$stem$+$ \textipa{\ae } & UR
\\ \midrule
\textipa{aamu} & \textipa{aamua} & \textipa{aamu}\\
\textipa{hopea} & \textipa{hopeaa} & \textipa{hopea}\\
\textipa{katto} & \textipa{kattoa} & \textipa{katto}\\
\textipa{kello} & \textipa{kelloa} & \textipa{kello}\\
\textipa{k\"ulm\ae } & \textipa{k\"ulm\ae \ae } & \textipa{k\"ulm\ae }\\
\textipa{koulu} & \textipa{koulua} & \textipa{koulu}\\
\textipa{lintu} & \textipa{lintua} & \textipa{lintu}\\
\textipa{h\"ull\"u} & \textipa{h\"ull\"u\ae } & \textipa{h\"ull\"u}\\
\textipa{k\"ompel\"o} & \textipa{k\"ompel\"o\ae } & \textipa{k\"ompel\"o}\\
\textipa{n\ae k\"o} & \textipa{n\ae k\"o\ae } & \textipa{n\ae k\"o}\\
\textipa{joki} & \textipa{jokea} & \textipa{joke}\\
\textipa{kivi} & \textipa{kive\ae } & \textipa{kive}\\
\textipa{muuri} & \textipa{muuria} & \textipa{muuri}\\
\textipa{naapuri} & \textipa{naapuria} &  -- \\
\textipa{nimi} & \textipa{nime\ae } &  -- \\
\textipa{kiirehti} & \textipa{kiirehti\ae } &  -- \\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[+low]| $\to$ \verb|[+back]| / \verb|[+rounded +back]| \verb|[ ]|* \verb|_| \\\textipa{e} $\to$ \verb|[+high]| /  \verb|_| \#
\end{tabular}

\pagebreak

\emph{Lithuanian:}\\\begin{longtable}{l|l}\toprule
\textipa{at} $+$stem$+$ \textipa{ti} & UR
\\ \midrule
\textipa{ateiti} & \textipa{ei}\\
\textipa{atimti} & \textipa{im}\\
\textipa{atleisti} & \textipa{leis}\\
\textipa{atlikti} & \textipa{lik}\\
\textipa{atko:pti} & \textipa{ko:p}\\
\textipa{atkurti} & \textipa{kur}\\
\textipa{adbekti} & \textipa{bek}\\
\textipa{adgauti} & \textipa{gau}\\
\textipa{adbukti} & \textipa{buk}\\
\textipa{adgimti} & \textipa{gim}\\
\textipa{atiduoti} & \textipa{iduo}\\
\textipa{atidari:ti} & \textipa{idari:}\\
\textipa{atideti} &  -- \\
\textipa{atiteisti} &  -- \\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[-vowel]| $\to$ \textipa{d} /  \verb|_| \verb|[-sonorant +voice]|
\end{tabular}

\pagebreak

\emph{Kerewe:}\\\begin{longtable}{llllllll|l}\toprule
\textipa{ku} $+$stem$+$ \textipa{a} & \textipa{ku} $+$stem$+$ \textipa{ana} & \textipa{ku} $+$stem$+$ \textipa{ila} & \textipa{ku} $+$stem$+$ \textipa{ilana} & \textipa{kut\'{u}} $+$stem$+$ \textipa{a} & \textipa{kuk\'{i}} $+$stem$+$ \textipa{a} & \textipa{kut\'{u}} $+$stem$+$ \textipa{ila} & \textipa{kuk\'{i}t\'{u}} $+$stem$+$ \textipa{ila} & UR
\\ \midrule
\textipa{kubala} & \textipa{kubalana} & \textipa{kubalila} & \textipa{kubalilana} & \textipa{kut\'{u}b\'{a}la} & \textipa{kuk\'{i}b\'{a}la} & \textipa{kut\'{u}b\'{a}lila} & \textipa{kuk\'{i}t\'{u}balila} & \textipa{bal}\\
\textipa{kugaja} & \textipa{kugajana} & \textipa{kugajila} & \textipa{kugajilana} & \textipa{kut\'{u}g\'{a}ja} & \textipa{kuk\'{i}g\'{a}ja} & \textipa{kut\'{u}g\'{a}jila} & \textipa{kuk\'{i}t\'{u}gajila} & \textipa{gaj}\\
\textipa{kugula} & \textipa{kugulana} & \textipa{kugulila} & \textipa{kugulilana} & \textipa{kut\'{u}g\'{u}la} & \textipa{kuk\'{i}g\'{u}la} & \textipa{kut\'{u}g\'{u}lila} & \textipa{kuk\'{i}t\'{u}gulila} & \textipa{gul}\\
\textipa{kub\'{a}la} & \textipa{kub\'{a}l\'{a}na} & \textipa{kub\'{a}l\'{i}la} & \textipa{kub\'{a}l\'{i}lana} & \textipa{kut\'{u}b\'{a}la} & \textipa{kuk\'{i}b\'{a}la} & \textipa{kut\'{u}b\'{a}lila} & \textipa{kuk\'{i}t\'{u}balila} & \textipa{b\'{a}l}\\
\textipa{kul\'{u}ma} & \textipa{kul\'{u}m\'{a}na} & \textipa{kul\'{u}m\'{i}la} & \textipa{kul\'{u}m\'{i}lana} & \textipa{kut\'{u}l\'{u}ma} & \textipa{kuk\'{i}l\'{u}ma} & \textipa{kut\'{u}l\'{u}mila} & \textipa{kuk\'{i}t\'{u}lumila} &  -- \\
\textipa{kus\'{u}na} & \textipa{kus\'{u}n\'{a}na} & \textipa{kus\'{u}n\'{i}la} & \textipa{kus\'{u}n\'{i}lana} & \textipa{kut\'{u}s\'{u}na} & \textipa{kuk\'{i}s\'{u}na} & \textipa{kut\'{u}s\'{u}nila} & \textipa{kuk\'{i}t\'{u}sunila} &  -- \\
\textipa{kul\'{a}ba} & \textipa{kul\'{a}b\'{a}na} & \textipa{kul\'{a}b\'{i}la} & \textipa{kul\'{a}b\'{i}lana} & \textipa{kut\'{u}l\'{a}ba} & \textipa{kuk\'{i}l\'{a}ba} & \textipa{kut\'{u}l\'{a}bila} & \textipa{kuk\'{i}t\'{u}labila} &  -- \\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[ ]| $\to$ \verb|[-highTone]| / \verb|[+highTone]| \verb|[ ]| \verb|_| \\\verb|[ ]| $\to$ \verb|[+highTone]| / \verb|[+highTone]| \verb|[ ]| \verb|_| \verb|[-vowel]| \verb|[+vowel]|
\end{tabular}

\pagebreak

\emph{Polish:}\\\begin{longtable}{ll|l}\toprule
$\varnothing$ $+$stem$+$ $\varnothing$ & $\varnothing$ $+$stem$+$ \textipa{i} & UR
\\ \midrule
\textipa{klup} & \textipa{klubi} & \textipa{klub}\\
\textipa{trup} & \textipa{trupi} & \textipa{trup}\\
\textipa{dom} & \textipa{domi} & \textipa{dom}\\
\textipa{snop} & \textipa{snopi} & \textipa{snp}\\
\textipa{Zwup} & \textipa{Zwobi} & \textipa{Zwob}\\
\textipa{trut} & \textipa{trudi} & \textipa{trud}\\
\textipa{dzvon} & \textipa{dzvoni} & \textipa{dzvon}\\
\textipa{kot} & \textipa{koti} & \textipa{kt}\\
\textipa{lut} & \textipa{lodi} & \textipa{lod}\\
\textipa{grus} & \textipa{gruzi} & \textipa{gruz}\\
\textipa{nos} & \textipa{nosi} & \textipa{ns}\\
\textipa{vus} & \textipa{vozi} & \textipa{voz}\\
\textipa{wuk} & \textipa{wugi} & \textipa{wug}\\
\textipa{wuk} & \textipa{wuki} & \textipa{wuk}\\
\textipa{sok} & \textipa{soki} & \textipa{sk}\\
\textipa{ruk} & \textipa{rogi} & \textipa{rog}\\
\textipa{bur} & \textipa{bori} & \textipa{bor}\\
\textipa{vuw} & \textipa{vowi} &  -- \\
\textipa{sul} & \textipa{soli} &  -- \\
\textipa{buj} & \textipa{boji} &  -- \\
\textipa{Sum} & \textipa{Sumi} &  -- \\
\textipa{Zur} & \textipa{Zuri} &  -- \\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[-sonorant]| $\to$ \verb|[-voice]| /  \verb|_| \#\\\verb|[+vowel]| $\to$ \textipa{u} /  \verb|_| \verb|[-nasal]| \#\\$\varnothing$ $\to$ \textipa{o} / \verb|[-vowel]| \verb|_| \verb|[-voice]|
\end{tabular}

\pagebreak

\emph{Ancient Greek:}\\\begin{longtable}{llll|l}\toprule
$\varnothing$ $+$stem$+$ \textipa{s} & $\varnothing$ $+$stem$+$ \textipa{os} & $\varnothing$ $+$stem$+$ \textipa{i} & $\varnothing$ $+$stem$+$ \textipa{si} & UR
\\ \midrule
\textipa{hals} & \textipa{halos} & \textipa{hali} & \textipa{halsi} & \textipa{hal}\\
\textipa{ojs} & \textipa{ojos} & \textipa{oji} & \textipa{ojsi} & \textipa{oj}\\
\textipa{sus} & \textipa{suos} & \textipa{sui} & \textipa{susi} & \textipa{su}\\
\textipa{klo:ps} & \textipa{klo:pos} & \textipa{klo:pi} & \textipa{klo:psi} & \textipa{klo:p}\\
\textipa{p\super hle:ps} & \textipa{p\super hle:bos} & \textipa{p\super hle:bi} & \textipa{p\super hle:psi} & \textipa{p\super hle:b}\\
\textipa{kate:lips} & \textipa{kate:lip\super hos} & \textipa{kate:lip\super hi} & \textipa{kate:lipsi} & \textipa{kate:lip\super h}\\
\textipa{p\super hulaks} & \textipa{p\super hulakos} & \textipa{p\super hulaki} & \textipa{p\super hulaksi} & \textipa{p\super hulak}\\
\textipa{ajks} & \textipa{ajgos} & \textipa{ajgi} & \textipa{ajksi} & \textipa{ajg}\\
\textipa{salpiNks} & \textipa{salpiNgos} & \textipa{salpiNgi} & \textipa{salpiNksi} & \textipa{salpiNg}\\
\textipa{onuks} & \textipa{onuk\super hos} & \textipa{onuk\super hi} & \textipa{onuksi} & \textipa{onuk\super h}\\
\textipa{t\super he:s} & \textipa{t\super he:tos} & \textipa{t\super he:ti} & \textipa{t\super he:si} & \textipa{t\super he:t}\\
\textipa{k\super haris} & \textipa{k\super haritos} & \textipa{k\super hariti} & \textipa{k\super harisi} & \textipa{k\super harit}\\
\textipa{elpis} & \textipa{elpidos} & \textipa{elpidi} & \textipa{elpisi} & \textipa{elpid}\\
\textipa{korus} & \textipa{korut\super hos} & \textipa{korut\super hi} & \textipa{korusi} & \textipa{korut\super h}\\
\textipa{ri:s} & \textipa{ri:nos} & \textipa{ri:ni} & \textipa{ri:si} & \textipa{ri:n}\\
\textipa{delp\super hi:s} & \textipa{delp\super hi:nos} & \textipa{delp\super hi:ni} & \textipa{delp\super hi:si} & \textipa{delp\super hi:n}\\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[-lateral +coronal]| $\to$ $\varnothing$ /  \verb|_| \verb|[-vowel]|\\\verb|[-sonorant]| $\to$ \verb|[-aspirated -voice]| /  \verb|_| \verb|[-voice]|
\end{tabular}

\pagebreak

\emph{Catalan:}\\\begin{longtable}{ll|l}\toprule
$\varnothing$ $+$stem$+$ $\varnothing$ & $\varnothing$ $+$stem$+$ \textipa{@} & UR
\\ \midrule
\textipa{@kelj} & \textipa{@kelj@} & \textipa{@kelj}\\
\textipa{mal} & \textipa{mal@} & \textipa{mal}\\
\textipa{siBil} & \textipa{siBil@} & \textipa{siBil}\\
\textipa{@skerp} & \textipa{@skerp@} & \textipa{@skerp}\\
\textipa{Sop} & \textipa{Sop@} & \textipa{Sop}\\
\textipa{sEk} & \textipa{sEk@} & \textipa{sEk}\\
\textipa{@spEs} & \textipa{@spEs@} & \textipa{@spEs}\\
\textipa{gros} & \textipa{gros@} & \textipa{gros}\\
\textipa{baS} & \textipa{baS@} & \textipa{baS}\\
\textipa{koS} & \textipa{koS@} & \textipa{koS}\\
\textipa{tot} & \textipa{tot@} & \textipa{tot}\\
\textipa{brut} & \textipa{brut@} & \textipa{brut}\\
\textipa{pOk} & \textipa{pOk@} & \textipa{pOk}\\
\textipa{pr@sis} & \textipa{pr@siz@} & \textipa{pr@siz}\\
\textipa{fr@nses} & \textipa{fr@nsez@} & \textipa{fr@nsez}\\
\textipa{gris} & \textipa{griz@} & \textipa{griz}\\
\textipa{k@zat} & \textipa{k@zaD@} & \textipa{k@zad}\\
\textipa{bwit} & \textipa{bwiD@} & \textipa{bwid}\\
\textipa{rOt\super S} & \textipa{rOZ@} & \textipa{rOd\super Z}\\
\textipa{bot\super S} & \textipa{boZ@} & \textipa{bod\super Z}\\
\textipa{orp} & \textipa{orB@} & \textipa{orb}\\
\textipa{ljark} & \textipa{ljarG@} & \textipa{ljarg}\\
\textipa{sek} & \textipa{seG@} & \textipa{seg}\\
\textipa{f@Suk} & \textipa{f@SuG@} & \textipa{f@Sug}\\
\textipa{grok} & \textipa{groG@} & \textipa{grog}\\
\textipa{puruk} & \textipa{puruG@} & \textipa{purug}\\
\textipa{kandit} & \textipa{kandiD@} & \textipa{kandid}\\
\textipa{frEt} & \textipa{frED@} & \textipa{frEd}\\
\textipa{s@Gu} & \textipa{s@Gur@} & \textipa{s@Gur}\\
\textipa{du} & \textipa{dur@} & \textipa{dur}\\
\textipa{s@G@Do} & \textipa{s@G@Dor@} & \textipa{s@g@Dor}\\
\textipa{kla} & \textipa{klar@} & \textipa{klar}\\
\textipa{nu} & \textipa{nu@} & \textipa{nu}\\
\textipa{kru} & \textipa{kru@} & \textipa{kru}\\
\textipa{flO\~nd\super Zu} & \textipa{flO\~nd\super Z@} & \textipa{flO\~nd\super Zu}\\
\textipa{dropu} & \textipa{drop@} & \textipa{dropu}\\
\textipa{@gzakt@} & \textipa{@gzakt@} & \textipa{@gzakt@}\\
\textipa{@lBi} & \textipa{@lBin@} & \textipa{@lBin}\\
\textipa{sa} & \textipa{san@} & \textipa{san}\\
\textipa{pla} & \textipa{plan@} & \textipa{plan}\\
\textipa{bo} & \textipa{bon@} & \textipa{bon}\\
\textipa{s@rE} & \textipa{s@rEn@} & \textipa{s@rEn}\\
\textipa{suBlim} & \textipa{suBlim@} & \textipa{suBlim}\\
\textipa{al} & \textipa{alt@} & \textipa{alt}\\
\textipa{fOr} & \textipa{fOrt@} & \textipa{fOrt}\\
\textipa{kur} & \textipa{kurt@} & \textipa{kurt}\\
\textipa{sor} & \textipa{sorD@} & \textipa{sord}\\
\textipa{bEr} & \textipa{bErD@} & \textipa{bErd}\\
\textipa{san} & \textipa{sant@} & \textipa{sant}\\
\textipa{k@lEn} & \textipa{k@lEnt@} & \textipa{k@lEnt}\\
\textipa{prufun} & \textipa{prufund@} & \textipa{prufund}\\
\textipa{f@kun} & \textipa{f@kund@} & \textipa{f@kund}\\
\textipa{d@sen} & \textipa{d@sent@} & \textipa{d@sent}\\
\textipa{dulen} & \textipa{dulent@} & \textipa{dulent}\\
\textipa{@stuDian} & \textipa{@stuDiant@} & \textipa{@stuDEiant}\\
\textipa{blaN} & \textipa{blaNk@} & \textipa{blaN\~n}\\
\bottomrule\end{longtable}
\begin{tabular}{l}\emph{Rules: }\\
\verb|[-lateral +sonorant +coronal]| $\to$ $\varnothing$ /  \verb|_| \#\\\verb|[-sonorant]| $\to$ \verb|[-voice]| /  \verb|_| \#\\\verb|[-vowel]| $\to$ \textipa{k} / \textipa{N} \verb|_| \\\verb|[-nasal +voice]| $\to$ \verb|[+continuant]| / \verb|[-nasal]| \verb|_| \verb|[+vowel]|\\\verb|[+vowel]| $\to$ $\varnothing$ / \verb|[-sonorant]| \verb|_| \verb|[+vowel]|\\\textipa{t} $\to$ $\varnothing$ / \verb|[-vowel]| \verb|_| \#
\end{tabular}

\end{document}