### 2018-06-16 manuscript.tex
science paper progress

% Use only LaTeX2e, calling the article.cls class and 12-point type.

\documentclass[12pt]{article}

% Users of the {thebibliography} environment or BibTeX should use the
% scicite.sty package, downloadable from *Science* at
% http://www.sciencemag.org/authors/preparing-manuscripts-using-latex 
% This package should properly format in-text
% reference calls and reference-list numbers.

\usepackage{scicite}

\usepackage{times}

% The preamble here sets up a lot of new/revised commands and
% environments.  It's annoying, but please do *not* try to strip these
% out into a separate .sty file (which could lead to the loss of some
% information when we convert the file to other formats).  Instead, keep
% them in the preamble of your main LaTeX source file.

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{tipa}
\usepackage{graphicx}
\usepackage{booktabs}
% The following parameters seem to provide a reasonable page setup.

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

\usepackage{phonrule}

%The next command sets up an environment for the abstract to your paper.

\newenvironment{sciabstract}{%
\begin{quote} \bf}
{\end{quote}}

% Include your paper's title here

\title{Synthesizing Theories of Human Language with Bayesian Program Induction}

% Place the author information here.  Please hand-code the contact
% information and notecalls; do *not* use \footnote commands.  Let the
% author contact information appear immediately below the author names
% as shown.  We would also prefer that you don't change the type-size
% settings shown here.

\author
{John Smith,$^{1\ast}$ Jane Doe,$^{1}$ Joe Scientist$^{2}$\\
\\
\normalsize{$^{1}$Department of Chemistry, University of Wherever,}\\
\normalsize{An Unknown Address, Wherever, ST 00000, USA}\\
\normalsize{$^{2}$Another Unknown Address, Palookaville, ST 99999, USA}\\
\\
\normalsize{$^\ast$To whom correspondence should be addressed; E-mail:  jsmith@wherever.edu.}
}

% Include the date command, but leave its argument blank.

\date{}

%%%%%%%%%%%%%%%%% END OF PREAMBLE %%%%%%%%%%%%%%%%

\begin{document}

% Double-space the manuscript.

\baselineskip24pt

% Make the title.

\maketitle

% Place your abstract within the special {sciabstract} environment.

\begin{sciabstract}
\end{sciabstract}

% In setting up this template for *Science* papers, we've used both
% the \section* command and the \paragraph* command for topical
% divisions.  Which you use will of course depend on the type of paper
% you're writing.  Review Articles tend to have displayed headings, for
% which \section* is more appropriate; Research Articles, when they have
% formal topical divisions at all, tend to signal them with bold text
% that runs into the paragraph, for which \paragraph* is the right
% choice.  Either way, use the asterisk (*) modifier, as shown, to
% suppress numbering.

\section*{Introduction}

An age-old aspiration within artificial intelligence research is to build a machine that automates the scientific process~\cite{paul1990autonomous,langley1987scientific}, either by designing experiments~\cite{sparkes2010towards} or
synthesizing theories and models~\cite{schmidt2009distilling}.
However, gaining traction on the problem of theory induction requires
a spectrum of tractable theory synthesis problems:
while early work attempted to recapitulate
classic discoveries like Kepler's laws~\cite{BACON},
these results proved brittle and difficult to generalize.
Despite these early works,
theory induction has never attained the same level of success as
machine vision or natural language processing.
How can the artificial intelligence community get theory induction off the ground?
This is an especially difficult question,
because the current mainstream in machine intelligence focuses on
learning to solve a qualitatively different class of problems (e.g., prediction tasks like classification and regression),
whereas theory induction requires
synthesizing human-understandable causal models of real-world phenomena~\cite{pearl2009causality},
so that human scientists can understand and learn from the AI's outputs.

We propose theory induction research start with theories of human
language, and scope out a collection of problems and techniques for a
key module of natural language: \emph{morphophonology}, the
relationship between pronunciations of words and their meaning.
Acquiring the morphophonology of a language involve solving a basic
problem that confronts both linguists and children: given a collection
of utterances, together with aspects of their meaning, what is the
causal relationship between form and meaning?  Morphophonology, and
natural language more broadly, presents a tractable challenge to
theory induction, for three reasons.  First, every natural language is
a different theory induction problem, so there is a wide spectrum of
induction tasks to evaluate on.  Second, linguistics as a field
traditionally formalizes theories in computational terms, and so there
is a pre-existing body of representations and algorithms for the AI
researcher to draw upon.  Third, we know that acquiring theories of
language (e.g., grammars) is tractable, because children 
acquire language, and field linguists can be trained to build theories of grammar.

Our contribution to theory induction is
an algorithm for synthesizing
theories of natural language morphophonology.
Like the linguist,
the algorithm starts with
a collection of utterances 
annotated with their meanings,
and then constructs a
causal, interpretable model explaining how
those meanings gave rise to the utterances.
An important aspect of a scientific theory is
that it makes predictions; here,
the theory predicts pronunciations of unobserved words conditioned on their meaning.
We evaluate our algorithm on \# data sets spanning \# languages,
automatically finding theories that can model a wide swath of a core component of human natural language.

\begin{table}\centering
\begin{tabular}{lcl}
  \toprule
  Example data&Morphophonology\\\midrule
\begin{tabular}{ll}
1&\textipa{\|x{j}ig}\\
4&\textipa{\|x{s}i}\\
5&\textipa{Na}\\
9&\textipa{gu}\\
10&\textipa{\|x{j}u}\\
11 ($= 10 + 1$) & \textipa{\|x{j}ug\|x{j}ig}\\
14 ($= 10 + 4$) & \textipa{\|x{j}ub\|x{s}i}\\
15 ($= 10 + 5$) & \textipa{\|x{j}uNa}\\
19 ($= 10 + 9$) & \textipa{\|x{j}urgu}\\
40 ($= 4 + 10$) & \textipa{\|x{s}ib\|x{j}u}\\
50 ($= 5 + 10$) & \textipa{Nab\|x{j}u}\\
90 ($= 9 + 10$) & \textipa{gub\|x{j}u}
\end{tabular}
&
\begin{tabular}{l}
  \emph{Consonant cluster reduction:}\\
  \phonb{C}{$\varnothing$}{\#}{C}\\
  (\emph{Upon encountering a consonant (C),}\\
  \emph{ delete it ($\to\varnothing$) at the beginning of}\\
  \emph{ a word (\#\_) if followed by another}\\
  \emph{ consonant (\_C)})
  \end{tabular}

  \\
  \bottomrule  \end{tabular}
\caption{Tibetan count system.}
\label{Tibetan}
\end{table}

\begin{figure}
  \includegraphics[width = \textwidth]{naturalLanguageGenerativeModel.pdf}
  \caption{Theory learner induces grammars (blue) for a range of languages, given observed form/meaning pairs (pink). Grammars are expressed as programs drawn from a universal grammar, or programming language. }\label{generativeModel}
  \end{figure}

\section*{Discovering Theories by Synthesizing Programs}

We frame our approach as Bayesian Program Learning (BPL: see~\cite{lake2015human}),
where the agent explains a
set of utterance/meaning pairs $\left\{(u_n,m_n) \right\}_{n = 1}^N$
by inferring a theory $T$, which we model as a program.
Formalizing grammars (theories) as generative programs has a long history in linguistics~\cite{chomsky1968sound}.
Written as a probabilistic inference problem, the agent seeks the $T$ maximizing $\left[\prod_{n = 1}^N P(u_n,m_n | T) \right] P(T|UG)$,
where $UG$ is a ``universal grammar'' (\textbf{TIM}: what should I cite here?)
constraining the space of allowed theories and imparting an inductive bias over the space of grammars.
In this BPL setting we model $UG$ as a prior distribution over theories (programs).
To model the space of programs
use Context-Sensitive Rewrites,
a Turing-complete program representation,
but restrict the rewrites
in such a way as to make them equivalent to finite state transducers --- this insight comes from the computational linguistics literature~\cite{beesley2003finite}.
For morphophonology,
the relationship between meaning $m_n$ and utterance $u_n$
involves a latent \emph{lexicon},
which is a mapping between lexemes and the pronunciation of their stem.
For children, linguists, and our model, the lexicon
must also be inferred (Fig.~\ref{generativeModel}).

Although this framing captures the problem a BPL theory inductor needs to solve,
it offers no guidance on how to solve that problem:
the space of all programs (theories) is infinitely large and
sharply discontinuous, lacking the local smoothness that
enables local optimization algorithms (e.g., gradient descent; MCMC)
to succeed.
We adopt a strategy currently unconventional to the
AI community: constraint--based program synthesis,
where the optimization problem is translated
into a combinatorial constraint satisfaction 
problem and solved using
a SMT solver~\cite{solar2006combinatorial}.

\section*{Discussion}

\bibliography{main}

\bibliographystyle{Science}

%% \section*{Acknowledgments}
%% Include acknowledgments of funding, any patents pending, where raw data for the paper are deposited, etc.

%% %Here you should list the contents of your Supplementary Materials -- below is an example. 
%% %You should include a list of Supplementary figures, Tables, and any references that appear only in the SM. 
%% %Note that the reference numbering continues from the main text to the SM.
%% % In the example below, Refs. 4-10 were cited only in the SM.     
%% \section*{Supplementary materials}
%% Materials and Methods\\
%% Supplementary Text\\
%% Figs. S1 to S3\\
%% Tables S1 to S4\\
%% References \textit{(4-10)}

%% % For your review copy (i.e., the file you initially send in for
%% % evaluation), you can use the {figure} environment and the
%% % \includegraphics command to stream your figures into the text, placing
%% % all figures at the end.  For the final, revised manuscript for
%% % acceptance and production, however, PostScript or other graphics
%% % should not be streamed into your compliled file.  Instead, set
%% % captions as simple paragraphs (with a \noindent tag), setting them
%% % off from the rest of the text with a \clearpage as shown  below, and
%% % submit figures as separate files according to the Art Department's
%% % instructions.

%% \clearpage

%% \noindent {\bf Fig. 1.} Please do not use figure environments to set
%% up your figures in the final (post-peer-review) draft, do not include graphics in your
%% source code, and do not cite figures in the text using \LaTeX\
%% \verb+\ref+ commands.  Instead, simply refer to the figure numbers in
%% the text per {\it Science\/} style, and include the list of captions at
%% the end of the document, coded as ordinary paragraphs as shown in the
%% \texttt{scifile.tex} template file.  Your actual figure files should
%% be submitted separately.

\end{document}